\begin{dedication}
\vspace*{\fill}
\begin{center}
    To my parents, \\
    Steven and Penelope Smart.
\end{center}
\vspace*{\fill}
\end{dedication}


\begin{acknowledgements}
    The research presented here is not the result of my effort alone. Here I make my best attempt to acknowledge those whose impact was pivotal in bringing this research to fruition.

    First, and foremost, I would like to thank my advisor Prof.\ Yuan Ping. Her passion, wisdom, and creativity are unparalleled, I am extremely grateful for all that she has taught me in these five years. I am thankful to Dr.\ Feng Wu, who was a great mentor to me, his patience and excitement in sharing his knowledge was very inspiring to me. I would also like to thank Mingpeng Chen, who not only provided me incredible insight into experiment but also was an amazing friend to share a meal with. I would like to thank Kejun Li, who I worked on various two-dimensional material projects and Valentin Urena Baltazar, who I worked with on various transition metal oxide projects. To the entire Ping Group I am thankful for the many conversations we had in which I learned so much.

    Next, I thank several collaborators outside of the Ping Group. Dr.\ Tuan Anh Pham, was an amazing mentor to me, going beyond just research advice and often providing me with invaluable life and career advice. Dr.\ Tadashi Ogitsu and him had a huge impact in my PhD, giving me the opportunity to work at LLNL.\@ I am thankful to Dr.\ Marco Govoni at Argonne National Laboratory for facilitating the opportunity to be a part of the Quantum Information Science and Engineering Network. I am thankful to several experimentalist whose discussions and collaborations with I learned a lot from including, but not limited to, Prof.\ Kyoung-Shin Choi, Prof.\ Yat Li, Prof.\ Shaowei Chen, Prof.\ Frank `Bud' Bridges, and also their graduate students such as, Dr.\ Allison Cardiel, Dr.\ Tianyi Kou, Dr.\ Bingzhang Lu.

    I am thankful to my committee members, Prof.\ Frank `Bud' Bridges, Prof.\ Yat Li, and Prof.\ Yuan Ping, in particular Bud for agreeing to be the chair and my department advisor. There are also many members of the physics department who were invaluable to my PhD experience. Specifically, I am thankful to Dr.\ Ben Miller, Dr.\ Katie Hellier, and Prof.\ Stephano Profumo who provided me with a much needed guidance during my PhD. I thank Carey Williams and all the students in my year, for the fun times and terrific friendships.

    I am thankful to the physics department faculty at Humboldt State University who taught me physics such as Prof.\ Monty Mola, Prof.\ Wes Blivens, and Prof.\ C.D.\ Hoyle. I still remember the day I decided I wanted to major in Physics from a short discussion with Monty.

    Lastly, I am thankful to my family. My mother and father, who taught me that trying your best is the best way to enjoy life. My wife, Channey for being my best friend, and the ultimate source of wisdom and support. And my son Alden, who made life enjoyable and fun, even as I finished out the PhD during the coronavirus pandemic.

    % And thank you to the voice who told me that ``nobody enjoys their PhD'', this sentiment lit a scorching fire inside of me and proving it wrong was my greatest success.

    \hfill -TJS :)
\end{acknowledgements}
