\documentclass[12pt]{../include/ucthesis}
\input{../include/macros}

\begin{document}

\title{First-Principles Computational Design of Materials: From energy conversion to quantum information}
\author{Tyler J. Smart}
\degreeyear{2021}
\degreemonth{June}
\degree{DOCTOR OF PHILOSOPHY}
\chair{Prof.\ Frank Bridges}
\committeememberone{Prof.\ Yuan Ping}
\committeemembertwo{Prof.\ Yat Li}
%\committeememberthree{Professor 4} % Uncomment if you have 4 committee members. Also change the next line to `\numberofmembers{4}`
\numberofmembers{3}
\deanlineone{Quentin Williams}
\deanlinetwo{Interim Vice Provost and Dean of Graduate Studies}
\deanlinethree{}
\field{Physics}
\campus{Santa Cruz}

\begin{frontmatter}
\maketitle

\copyrightpage

\tableofcontents

\listoffigures

\listoftables

\begin{abstract}
    State-of-the-art first-principles calculations are implemented and utilized in order to optimize or give insight into material properties for various techonologies and applications. These applications, and therefore the research, have two primary focuses. The first is energy conversion and storage using transition metal oxides and the second is quantum defects in two-dimensional materials. Regarding the first, the limitation of transition metal oxides is in their poor carrier conduction due to the formation of small polarons. Here, detail and insight are presented into the formation of small polarons, especially under atomic doping wherein doping can help to improve carrier conduction in these systems and offer better energy conversion. For the second application, an emphasis in the design of single photon emitters and spin-based qubits in hexagonal boron nitride is presented. In particular, defect ionization energies, optoelectronic properties, and nonradiative propreties, including intersystem crossing, are discussed. Finally, selected topics are presented which do not necessarily fall into these two categories but is otherwise interesting to discuss.
\end{abstract}

\begin{dedication}
\vspace*{\fill}
\begin{center}
    To my wife, \\
    Channey Lane Smart.
\end{center}
\vspace*{\fill}
\end{dedication}

\begin{acknowledgements}
    The research presented here is not the result of my effort alone. Here I make my best attempt to acknowledge those whose impact was pivotal in bringing this research to fruition.

    First, and foremost, I would like to thank my advisor Prof.\ Yuan Ping. Her passion, wisdom, and creativity is unparalleled, I am extremely grateful for all that she has taught me in these five years. I am thankful to Dr.\ Feng Wu, who was an amazing mentor to me, his patience and excitement in sharing his knowledge was very inspiring to me. I would also like to thank Mingpeng Chen, who not only provided me incredible insight into experiment but also was an amazing friend to share a meal with. I would like to thank Kejun Li, who I worked on various two-dimensional material projects and Valentin Urena Baltazar, who I worked with on various transition metal oxide projects. To the entire Ping Group I am thankful for the many conversations we had in which I learned so much.

    Next, I thank several collaborators outside of the Ping Group. Dr.\ Tuan Anh Pham, was an amazing mentor to me, going beyond just research advice and often providing me with invaluable life and career advice. I thank him and Dr.\ Tadashi Ogitsu both for providing me the opportunity to collaborate with LLNL and learn so much from them. I am thankful Dr.\ Marco Govoni at Argonne National Laboratory for facilitating the opportunity to be apart of the Quantum Information Science and Engineering Network. I am thankful of several experimentalist whose discussions and collaborations with I learned a lot from including, but not limited to, Prof.\ Kyoung-Shin Choi, Prof.\ Yat Li, Prof.\ Shaowei Chen, Prof.\ Frank `Bud' Bridges, and also their graduate students such as, Dr.\ Allison Cardiel, Dr.\ Tianyi Kou, Dr.\ Bingzhang Lu.

    I am thankful to my committee members, Prof.\ Frank `Bud' Bridges, Prof.\ Yat Li, and Prof.\ Yuan Ping, in particular Bud for agreeing to be the chair and my department advisor. There are also many members of the physics department whose friendship and insights were invaluable to my PhD experience. Specifically, I am thankful to Dr.\ Ben Miller, Dr.\ Katie Hellier, and Prof.\ Stephano Profumo who provided me with a much needed guidance during my PhD. I thank Carey Williams and all the students in my year, for the fun times and a terrific friendship.

    I am thankful to physics department faculty at Humboldt State University who taught me physics such as Prof.\ Monty Mola, Prof.\ Wes Blivens, and Prof.\ C.D.\ Hoyle. I still remember the day I decided I wanted to major in Physics from a short discussion with Monty.

    Lastly, I am thankful to my family. My mother and father, who taught me that trying your best is the best way to enjoy life. My wife, Channey for being my best friend, and the ultimate source of wisdom and support. And my son Alden, who made life enjoyable and fun, even as I finished out the PhD during the coronavirus pandemic.

    Oh, and thank you to the voice who told me that ``nobody enjoys their PhD'', proving you wrong lit a scorching fire inside of me and was my greatest success. :)
\end{acknowledgements}

\end{frontmatter}

% ------------------------------------------------------------
% intro
\include{1-intro/intro}


% ------------------------------------------------------------
% tmo
\include{2-tmo/overview}
\include{2-tmo/fe2o3-hop}
\include{2-tmo/fe2o3-conc}
\include{2-tmo/fe2o3-clust}
\include{2-tmo/cuo}
\include{2-tmo/co3o4}
\include{2-tmo/exp}

% ------------------------------------------------------------
% 2d
\include{3-2d/overview}
\include{3-2d/cfe}
\include{3-2d/nonrad}
\include{3-2d/spin}

% ------------------------------------------------------------
% other
\chapter{Selected Topics}
\include{4-other/cat}
\include{4-other/2denv}
\include{4-other/nv}
\include{4-other/nerd}
\include{4-other/open}

% ------------------------------------------------------------
\chapter{Conclusions}
blank

\appendix

\include{5-app/nonrad}
\include{5-app/hopping}
\include{5-app/zfs}
\include{5-app/pl}
\include{5-app/optical}

% molecular orbital
% jahn teller
% codes

% \include{a2-pw2py}


\nocite{*}
% \bibliographystyle{plain}
\bibliographystyle{ieeetr}
\singlespacing
\bibliography{ref}

\doublespacing

\end{document}
